Emergent dual scaling of riverine biodiversity

### 2021-09-14 manuscript_pnas.Rmd
edits: minor text edit

We developed a theoretical framework synthesizing the influences of ecosystem size and complexity on biodiversity patterns in branching ecosystems. We depicted branching ecosystems as a spatial network of connected habitat patches which local communities inhabit [@yeakelSynchronisationStabilityRiver2014; @teruiMetapopulationStabilityBranching2018] (**Figure 1A and B**). In this network, the number of habitat patches $N_p$ (a theoretical proxy for ecosystem size $A$) and branching probability $P_b$ (ecosystem complexity) shape the spatial structure (**Materials and Methods**). We introduced two factors determining the temporal mean of environmental condition at each habitat patch $\mu_z$: (1) headwater environments (the most upstream habitat patch) and (2) local environmental noise. Environmental values at headwaters are drawn randomly from a normal distribution with a mean of zero and a standard deviation of $\sigma_h$ and propagate downstream. We modeled the downstream propagation as a spatial autoregressive process, in which local noise is added to the adjacent upstream environment as $\mu_{z,k} \sim Normal(\mu_{z,k+1}, \sigma_{l}^2)$ ($k$ is the network distance from the outlet patch; $k = 1$ at the outlet). The environmental values recurrently 'mix' at confluences considering the relative size of joining tributaries (see **Materials and Methods**). Therefore, our network-generation procedure resembles natural processes of how branching river networks create diverse habitats in a metacommunity [@mcguireNetworkAnalysisReveals2014].

>>>>

We developed a theoretical framework synthesizing the influences of ecosystem size and complexity on biodiversity patterns in branching ecosystems. We depicted branching ecosystems as a spatial network of connected habitat patches which local communities inhabit [@yeakelSynchronisationStabilityRiver2014; @teruiMetapopulationStabilityBranching2018] (**Figure 1A and B**). In this network, the number of habitat patches $N_p$ (a theoretical proxy for ecosystem size $A$) and branching probability $P_b$ (ecosystem complexity) shape the spatial structure (**Materials and Methods**). We introduced two factors determining the temporal mean of environmental condition at each habitat patch $\mu_z$: headwater environments (the most upstream habitat patch) and local environmental noise. Environmental values at headwaters are drawn randomly from a normal distribution with a mean of zero and a standard deviation of $\sigma_h$ and propagate downstream. We modeled the downstream propagation as a spatial autoregressive process, in which local noise is added to the adjacent upstream environment as $\mu_{z,k} \sim Normal(\mu_{z,k+1}, \sigma_{l}^2)$ ($k$ is the network distance from the outlet patch; $k = 1$ at the outlet). The environmental values recurrently 'mix' at confluences considering the relative size of joining tributaries (see **Materials and Methods**). Therefore, our network-generation procedure resembles natural processes of how branching river networks create diverse habitats in a metacommunity [@mcguireNetworkAnalysisReveals2014].
